Data Cleaning logbook

### 2018-12-07 docs/manuscript.Rmd
Updated discussion to talk about lessons learned

While the evaluated reserves have failed to provide fishery benefits up to now, there are a number of additional ecological, fisheries, and social benefits. Marine reserves provide protection to a wider range of species and vulnerable habitat. These sites can serve as an insurance against uncertainty and errors in fisheries management, as well as mild environmental shocks \citep{micheli_2012-EU,deleo_2015,roberts_2017-J9,aalto}. Self-regulation of fishing effort (*i.e.* reduction in harvest) can serve as a way to compensate for future declines associated to environmental variation \citep{finkbeiner_2018}. Embarking in a marine conservation project can bring the community together, which promotes social cohesion and builds social capital \citep{fulton_2019}. Showing commitment to marine conservation and sustainable fishing practices allows fishers to have greater bargaining power and leverage over fisheries management \citep{prezramrez_2012}. Furthermore, our work evaluates only TURF-reserves established as Fish Refuges and is not representative of other Fish Refuges not implemented within TURFs. Future research should aim at evaluating other areas that have also been established as bottom-up processes but without the presence of TURFs (*e.g.* \citet{dof_websiteC_2012}), others established through a top-down process (*i.e.* \citet{dof_websiteU_2018}), as well as the relationship between governance and effectiveness across this gradient of approaches.

Community-based marine reserves in small-scale fisheries can be helpful conservation and fishery management tools when appropriately implemented. Lessons learned from these cases can guide implementation of community-based marine reserves elsewhere. For the particular case of the marine reserves that we evaluate, the possibility of expanding reserves or merging existing polygons into larger areas should be evaluated and proposed to the communities. Community-based marine reserves might have more benefits that result from indirect effects of the reserves, particularly providing resilience to shocks and management errors, and promoting social cohesion, which should be taken into account when evaluating the outcomes of TURF-reserves. Having full community support surely represents an advantage, but it is important that community-based TURF-reserves meet essential design principles such as size and placement so as to maximize their effectiveness.

>>>>

While the evaluated reserves have failed to provide fishery benefits up to now, there are a number of additional ecological, fisheries, and social benefits. Marine reserves provide protection to a wider range of species and vulnerable habitat. These sites can serve as an insurance against uncertainty and errors in fisheries management, as well as mild environmental shocks \citep{micheli_2012-EU,deleo_2015,roberts_2017-J9,aalto}. Self-regulation of fishing effort (*i.e.* reduction in harvest) can serve as a way to compensate for future declines associated to environmental variation \citep{finkbeiner_2018}. Embarking in a marine conservation project can bring the community together, which promotes social cohesion and builds social capital \citep{fulton_2019}. Showing commitment to marine conservation and sustainable fishing practices allows fishers to have greater bargaining power and leverage over fisheries management \citep{prezramrez_2012}. Furthermore, our work evaluates only TURF-reserves established as Fish Refuges and is not representative of other Fish Refuges implemented without TURFs. Future research should aim at evaluating other areas that have also been established as bottom-up processes but without the presence of TURFs (*e.g.* \citet{dof_websiteC_2012}), others established through a top-down process (*i.e.* \citet{dof_websiteU_2018}), as well as the relationship between governance and effectiveness across this gradient of approaches.

Community-based TURF-reserves in small-scale fisheries can be helpful conservation and fishery management tools when appropriately implemented. As we move forward in marine conservation, we must recognize the value of local knowledge to identify priority areas for conservation, but without setting design principles aside. Lessons learned from these cases can guide implementation of community-based TURF-reserves elsewhere. For the particular case of the ones we evaluate, the possibility of expanding reserves or merging existing polygons into larger areas should be evaluated and proposed to the communities. Community-based marine reserves might have more benefits that result from indirect effects of the reserves, particularly providing resilience to shocks and management errors, and promoting social cohesion, which should be taken into account when evaluating the outcomes of TURF-reserves. Having full community support surely represents an advantage, but it is important that community-based TURF-reserves meet essential design principles such as size and placement so as to maximize their effectiveness.

### 2018-12-07 docs/manuscript.tex
Updated discussion to talk about lessons learned

The biological data are collectged by members from each community and
personnel from the Mexican CSO \emph{Comunidad y Biodiversidad}
(\href{www.cobi.org.mx}{COBI}). Trained divers record richness and
abundances of fish and invertebrate species along replicate transects
(30 \(\times\) 2 m each) at depths 5-20 m in the reserves and control
sites \citep{fulton_2018,fulton_2019,suman_2010-ez}. Size structures are
also collected during fish surveys. All sites were surveyed annually,
and at least once before implementation of the reserves. A summary of
sampling effort (number of transects per site) is shown in the
supplementary materials (S1 Table and S2 Table).

@@

Indicators showed ambiguous responses through time for each reserve.
Figure \ref{fig:indicators}A shows positive effect sizes for lobster
densities in Isla Natividad and Punta Herrero during the first years,
but the effect is eroded through time. In the case of Maria Elena,
positive changes were observed in the third and fourth year. These
effects are in the order of 0.2 extra organisms \(\mathrm{m}^{-2}\) for
Isla Natividad and Punta Herrero, and 0.01 organisms \(\mathrm{m}^{-2}\)
for Maria Elena, but are not significantly different from zero
(\(p > 0.05\)). Likewise, no significant changes were detected in fish
biomass or invertebrate and fish densities (Fig.
\ref{fig:indicators}B-D), where effect sizes oscillated around zero
without clear trends. Full tables with model coefficients are presented
in the supplementary materials \textbf{(S1 Table, S2 Table, S3 Table)}.

>>>>

The biological data are collectged by members from each community and
personnel from the Mexican CSO \emph{Comunidad y Biodiversidad}
(\href{www.cobi.org.mx}{COBI}). Trained divers record richness and
abundances of fish and invertebrate species along replicate transects
(30 \(\times\) 2 m each) at depths 5-20 m in the reserves and control
sites \citep{fulton_2018,fulton_2019,suman_2010-ez}. Size structures are
also collected during fish surveys. All sites were surveyed annually,
and at least once before implementation of the reserves. A summary of
sampling effort (number of transects per site) is shown in the
supplementary materials (Tables S1-S2).

@@

Indicators showed ambiguous responses through time for each reserve.
Figure \ref{fig:indicators}A shows positive effect sizes for lobster
densities in Isla Natividad and Punta Herrero during the first years,
but the effect is eroded through time. In the case of Maria Elena,
positive changes were observed in the third and fourth year. These
effects are in the order of 0.2 extra organisms \(\mathrm{m}^{-2}\) for
Isla Natividad and Punta Herrero, and 0.01 organisms \(\mathrm{m}^{-2}\)
for Maria Elena, but are not significantly different from zero
(\(p > 0.05\)). Likewise, no significant changes were detected in fish
biomass or invertebrate and fish densities (Fig.
\ref{fig:indicators}B-D), where effect sizes oscillated around zero
without clear trends. Figures and tables with timeseries of indicators
and model coefficients are presented in the supplementary materials
(Figures S1-S2, Tables S3-S5).

### 2018-12-07 scripts/data_cleaning_logbook.Rmd
Updated discussion to talk about lessons learned

@@

```{r}
species_bio <- MPAtools::species_bio %>% 
  mutate(Genus = as.character(str_split(GeneroEspecie, pattern = " ", simplify = T)[,1]),
         Genus = str_replace(Genus, " ", ""))

needed_spp <- c("Kyphosus incisor",
                "Kyphosus sectatrix",
                "Pterois volitans",
                "Scarus guacamaia",
                "Chaetodon capistratus",
                "Chaetodon capistratus",
                "Chaetodon striatus",
                "Calamus pennatula")

@@

@@

GenusLevelBio <- species_bio %>% 
  filter(Genus %in% needed_genus) %>%
  select(GeneroEspecie = Genus, a, b) %>% 
  group_by(GeneroEspecie) %>% 
  summarize_all(mean, na.rm = T) %>% 
  ungroup()

species_bio %<>%
  select(GeneroEspecie, a, b) %>% 
  rbind(FB_update) %>% 
  rbind(GenusLevelBio) %>% 
  janitor::clean_names() %>% 
    rename(generoespecie = genero_especie)
```

@@

```{r}
fish <- read.csv(here::here("raw_data", "fish_data.csv"), strip.white = T, stringsAsFactors = F) %>% 
  janitor::clean_names() %>% 
  rename(generoespecie = genero_especie) %>% 
  filter(comunidad %in% c("Isla Natividad", "Maria Elena", "Punta Herrero")) %>% 
  filter(!is.na(rc),
         rc %in% c("La Plana / Las Cuevas - La Dulce / Babencho",
                   "Cabezo - Cabezo (Control)",
                   "Manchon - Manchon (Control)",
                   "El Faro - El Faro (Control)")) %>% 
  filter(!(rc %in% c("El Faro - El Faro (Control)") & ano == 2012)) %>% 
  filter(abundancia > 0) %>% 
  mutate(generoespecie = case_when(generoespecie == "Cephalopholis cruentatus" ~ "Cephalopholis cruentata",
                                   generoespecie == "Haemulon macrostomun" ~ "Haemulon macrostomum",
                                   generoespecie == "Xcochin spp" ~ "Balistes capriscus",
                                   generoespecie == "Pterois volitans/miles" ~ "Pterois volitans",
                                   generoespecie == "Mycteroperca spp" ~ "Mycteroperca",
                                   generoespecie == "Chromis spp" ~ "Chromis",
                                   generoespecie == "Damisela spp" ~ "Stegastes",
                                   generoespecie == "Haemulon spp" ~ "Haemulon",
                                   generoespecie == "Lancero spp" ~ "Acanthurus",
                                   generoespecie == "Kyphosus sectatrix/incisor" ~ "Kyphosus",
                                   generoespecie == "Sebastes spp" ~ "Sebastes",
                                   generoespecie == "Loro spp" ~ "Scarus",
                                   generoespecie == "Lutjanus spp" ~ "Lutjanus",
                                   generoespecie == "Mariposa spp" ~ "Chaetodon",
                                   generoespecie == "Pluma spp" ~ "Calamus pennatula",
                                   generoespecie == "Sparisoma aurolineatum" ~ "Sparisoma aurofrenatum",
                                   generoespecie == "Vieja spp" ~ "Bodianus rufus",
                                   TRUE ~ generoespecie)) %>% 
  left_join(species_bio, by = c("generoespecie")) %>% 
  group_by(comunidad, ano, rc, sitio, zona, transecto, generoespecie, talla, a,b) %>% 
  summarize(abundancia = sum(abundancia)) %>% 
  ungroup() %>% 
  mutate(year = case_when(comunidad == "Isla Natividad" ~ ano - 2006,
                         comunidad == "Maria Elena" ~ ano - 2012,
                         TRUE ~ ano - 2013)) %>% 
  select(comunidad, ano, year, rc, sitio, zona, transecto, generoespecie, talla, abundancia, a, b) %>% 
  arrange() %>% 
  drop_na(a, b)
```

@@

```{r}
conapesca <- readRDS(here::here("raw_data","conapesca.rds")) %>% 
  mutate(UnidadEconomica = as.character(UnidadEconomica),
         LugarDeCaptura = as.character(LugarDeCaptura)) %>% 
  mutate(UnidadEconomica = ifelse(UnidadEconomica == "Scpp bahia tortugas s.c de r.l", 
                                  "Scpp bahia tortugas scl",
                                  UnidadEconomica)) %>% 
  filter(UnidadEconomica %in% cooperatives) %>%
  filter(NombreComun %in% c("Langosta roja ent. fca.", "Langosta ent. fca.")) %>%
  filter(!(UnidadEconomica == "Scpp cozumel scl" & !LugarDeCaptura %in% c("Bahia espiritu santo",
                                                                          "Bahia del del espiritu santo",
                                                                          "Bahia espiritu santo"))) %>% 
  group_by(UnidadEconomica, Ano, NombreComun) %>%
  summarize(PesoDesembarcado = sum(PesoDesembarcado), 
            PesoVivo = sum(PesoVivo),
            Valor = sum(Valor)) %>%
  ungroup() %>% 
  filter(!(UnidadEconomica == "Scpp buzos y pescadores de la baja california scl" & NombreComun == "Langosta ent. fca.")) %>% 
  left_join(cpi, by = "Ano") %>% 
  mutate(Valor = Valor * (CPI / cpi_2014),
         Group = case_when(UnidadEconomica %in% cooperatives_t ~ "Treated",
                           TRUE ~ "Control"),
         UnidadEconomica = gsub(pattern = "Scpp ", replacement = "", x = UnidadEconomica),
         UnidadEconomica = gsub(pattern = " scl", replacement = "", x = UnidadEconomica),
         UnidadEconomica = gsub(pattern = " s.c de r.l", replacement = "", x = UnidadEconomica),
         UnidadEconomica = gsub(pattern = " sc de rl", replacement = "", x = UnidadEconomica)) %>% 
  left_join(L_conapesca, by = "UnidadEconomica") %>% 
  mutate(Post = ifelse(Ano <= Start, 0, 1),
         years_centered = Ano - Start) %>% 
  select(Zona, Grupo = Group, Comunidad, UnidadEconomica, Ano, Post, years_centered, NombreComun, PesoVivo, PesoDesembarcado, Valor, CPI) %>% 
  janitor::clean_names()
```

# Export all data

>>>>

The commit history for this file can be seen [here](https://github.com/jcvdav/ReserveEffect/commits/master/scripts/data_cleaning_logbook.Rmd)

@@

```{r}
species_bio <- MPAtools::species_bio %>% 
  mutate(Genus = as.character(str_split(GeneroEspecie, pattern = " ", simplify = T)[,1]),
         Genus = str_replace(Genus, " ", "")) %>% 
  drop_na(a, b)

needed_spp <- c("Embiotoca jacksoni",
                "Kyphosus incisor",
                "Kyphosus sectatrix",
                "Pterois volitans",
                "Scarus guacamaia",
                "Chaetodon capistratus",
                "Chaetodon capistratus",
                "Chaetodon striatus",
                "Calamus pennatula")

@@

manual_fishbase_bayesian <- tibble(
  GeneroEspecie = c("Chromis punctipinnis",
                    "Oxyjulis californica",
                    "Heterodontus francisci",
                    "Rhacochilus vacca",
                "Squatina californica"),
  a = c(0.01514, 0.00447, 0.00355, 0.01950, 0.00741),
  b = c(2.99, 3.14,3.16, 3.02, 3.04)
)

@@

GenusLevelBio <- species_bio %>% 
  select(Genus, ag = a, bg = b) %>% 
  group_by(Genus) %>% 
  summarize_all(mean, na.rm = T) %>% 
  ungroup()

species_bio %<>%
  select(GeneroEspecie, a, b) %>% 
  rbind(FB_update) %>% 
  rbind(manual_fishbase_bayesian) %>% 
  janitor::clean_names() %>% 
    rename(generoespecie = genero_especie)
```

@@

```{r}
fish <- read.csv(here::here("raw_data", "fish_data.csv"), strip.white = T, stringsAsFactors = F) %>% 
  janitor::clean_names() %>% 
  rename(generoespecie = genero_especie) %>% 
  filter(comunidad %in% c("Isla Natividad", "Maria Elena", "Punta Herrero")) %>% 
  filter(!is.na(rc),
         rc %in% c("La Plana / Las Cuevas - La Dulce / Babencho",
                   "Cabezo - Cabezo (Control)",
                   "Manchon - Manchon (Control)",
                   "El Faro - El Faro (Control)")) %>% 
  filter(!(rc %in% c("El Faro - El Faro (Control)") & ano == 2012)) %>% 
  filter(abundancia > 0) %>% 
  mutate(generoespecie = case_when(generoespecie == "Cephalopholis cruentatus" ~ "Cephalopholis cruentata",
                                   generoespecie == "Haemulon macrostomun" ~ "Haemulon macrostomum",
                                   generoespecie == "Xcochin spp" ~ "Balistes capriscus",
                                   generoespecie == "Pterois volitans/miles" ~ "Pterois volitans",
                                   generoespecie == "Mycteroperca spp" ~ "Mycteroperca",
                                   generoespecie == "Chromis spp" ~ "Chromis",
                                   generoespecie == "Damisela spp" ~ "Stegastes",
                                   generoespecie == "Haemulon spp" ~ "Haemulon",
                                   generoespecie == "Lancero spp" ~ "Acanthurus",
                                   generoespecie == "Kyphosus sectatrix/incisor" ~ "Kyphosus",
                                   generoespecie == "Sebastes spp" ~ "Sebastes",
                                   generoespecie == "Loro spp" ~ "Scarus",
                                   generoespecie == "Lutjanus spp" ~ "Lutjanus",
                                   generoespecie == "Mariposa spp" ~ "Chaetodon",
                                   generoespecie == "Pluma spp" ~ "Calamus pennatula",
                                   generoespecie == "Sparisoma aurolineatum" ~ "Sparisoma aurofrenatum",
                                   generoespecie == "Vieja spp" ~ "Bodianus rufus",
                                   TRUE ~ generoespecie)) %>% 
  mutate(Genus = as.character(str_split(generoespecie, pattern = " ", simplify = T)[,1]),
         Genus = str_replace(Genus, " ", "")) %>% 
  left_join(species_bio, by = "generoespecie") %>% 
  left_join(GenusLevelBio, by = "Genus") %>%
  mutate(a = ifelse(is.na(a), ag, a),
         b = ifelse(is.na(b), bg, b)) %>% 
  group_by(comunidad, ano, rc, sitio, zona, transecto, generoespecie, talla, a,b) %>% 
  summarize(abundancia = sum(abundancia)) %>% 
  ungroup() %>% 
  mutate(year = case_when(comunidad == "Isla Natividad" ~ ano - 2006,
                         comunidad == "Maria Elena" ~ ano - 2012,
                         TRUE ~ ano - 2013))

fish <- fish %>% 
  select(comunidad, ano, year, rc, sitio, zona, transecto, generoespecie, talla, abundancia, a, b) %>% 
  arrange()
```

@@

```{r}
conapesca <- readRDS(here::here("raw_data","conapesca.rds")) %>% 
  mutate(UnidadEconomica = as.character(UnidadEconomica),
         LugarDeCaptura = as.character(LugarDeCaptura)) %>% 
  mutate(UnidadEconomica = ifelse(UnidadEconomica == "Scpp bahia tortugas s.c de r.l", 
                                  "Scpp bahia tortugas scl",
                                  UnidadEconomica)) %>% 
  filter(UnidadEconomica %in% cooperatives) %>%
  filter(NombreComun %in% c("Langosta roja ent. fca.", "Langosta ent. fca.")) %>%
  filter(!(UnidadEconomica == "Scpp cozumel scl" & !LugarDeCaptura %in% c("Bahia espiritu santo",
                                                                          "Bahia del del espiritu santo",
                                                                          "Bahia espiritu santo"))) %>% 
  group_by(UnidadEconomica, Ano, NombreComun) %>%
  summarize(PesoDesembarcado = sum(PesoDesembarcado), 
            PesoVivo = sum(PesoVivo),
            Valor = sum(Valor)) %>%
  ungroup() %>% 
  filter(!(UnidadEconomica == "Scpp buzos y pescadores de la baja california scl" & NombreComun == "Langosta ent. fca.")) %>% 
  left_join(cpi, by = "Ano") %>% 
  mutate(Valor = Valor * (CPI / cpi_2014),
         Group = case_when(UnidadEconomica %in% cooperatives_t ~ "Treated",
                           TRUE ~ "Control"),
         UnidadEconomica = gsub(pattern = "Scpp ", replacement = "", x = UnidadEconomica),
         UnidadEconomica = gsub(pattern = " scl", replacement = "", x = UnidadEconomica),
         UnidadEconomica = gsub(pattern = " s.c de r.l", replacement = "", x = UnidadEconomica),
         UnidadEconomica = gsub(pattern = " sc de rl", replacement = "", x = UnidadEconomica)) %>% 
  left_join(L_conapesca, by = "UnidadEconomica") %>% 
  mutate(Post = ifelse(Ano <= Start, 0, 1),
         years_centered = Ano - Start) %>% 
  select(Zona, Grupo = Group, Comunidad, UnidadEconomica, Ano, Post, years_centered, NombreComun, PesoVivo, PesoDesembarcado, Valor, CPI) %>% 
  janitor::clean_names() %>% 
  filter(!(zona == "MC" & years_centered %in% c(-3, 2))) %>% 
  filter(years_centered > -7)
```
